Week 1: Data Types
======

### Contents

* Data Types
  * [atomic classes](#classes): numeric, logical, character, integer, complex
  * [vectors](#vectors), [lists](#lists)
  * [factors](#factors)
  * [missing values](#missing-values)
  * [data frames](#data-frames)
  * [names](#names)

### Classes

* Five basic classes of objects:
  * character
  * numeric (real numbers) = Double
  * integer = Int
  * complex
  * logical (True/False) = Boolean

* Two basic object
  * vector: contain objects of the same class
  * list: contain objects of difference classes

* Numbers (numeric)
   * integer with `L` suffix
   * `Inf` represent infinity: `1 / 0` is `Inf`, `1 / Inf` is 0
   * `NaN` represents undefined value: `0 / 0` is `NaN`

* Attributes
  * names, dimnames
  * dimansions: matrices, arrays
  * class
  * length
  * other user-defined attributes/metadata
  * can be accessed using `attribute()`

* Entering Input

```{r}
msg <- "hello" ## assignment
```

* Evaluation

```{r}
x <- 5 ## nothing printed
x      ## auto-printing occurs
print(x) ## explicit printing
```

* Printing

```{r}
x <- 1:10 ## the : operator is used to create integer sequences
x
```

### Vectors

* Creating Vectors of objects using the `c()` (stands for concatenate) function

```{r}
x <- c(0.5, 0.6)      ## numeric
x <- c(TRUE, FALSE)   ## logical
x <- c(T, F)          ## logical
x <- c("a", "b", "c") ## character
x <- 0:10             ## integer
x <- c(1+0i, 2+4i)    ## complex
```

* Creating Vectors of objects using the `vector()` function with class defined as first parameter and length as second parameter

```{r}
x <- vector("numeric", length = 10)
x
```

* Mixing Objects: coercion happens to change less common object to more common object

```{r}
y <- c(1.7, "a")   ## character
y <- c(TRUE, 2)    ## numeric
y <- c("a", TRUE)  ## character
```

* Explicit Coercion by using `as.*`

```{r}
x <- 0:6
class(x)
as.numeric(x)
as.logical(x)
as.character(x)
```

* Explicit Coercion with nonsensical coercion

```{r}
x <- c("a", "b", "c")
as.numeric(x) ## get warning
as.logical(x)
as.complex(x)
```

### Matrices

* Matrices: vectors with a _dimension_ attribute. The _dimension_ attribute is itself an integer vector of length 2 `(nrow, ncol)`

```{r}
m <- matrix(nrow = 2, ncol = 3)
m
dim(m)
attributes(m)
```

* Matrices are constructed _column-wise_, so entries can be though of starting in the upper-left corner and running down the columns

```{r}
m <- matrix(1:6, nrow = 2, ncol = 3)
m
```

* Matrices can be created directly from vectors by adding a dimension attribute

```{r}
m <- 1:10
m
dim(m) <- c(2, 5)
m
```

* Matrices can be created by _column-binding_ or _row-binding_ with `cbind()` and `rbind()`

```{r}
x <- 1:3
y <- 10:12
cbind(x, y)
rbind(x, y)
```

### Lists

* Lists are a special type of vectors that can contain elements of different classes

```{r}
x <- list(1, "a", TRUE, 1 + 4i)
x ## with double brackets
```

### Factors

* Factors are used to represent categorical data, can be ordered or unordered. 
  * Factors are treated specially by modelling function like `lm()` and `glm()`
  * Using factors with labels is better than using integers becuase factors are self-describing; having a variable that has values "Male" and "Female" is better than a variable that has values 1 and 2
  
```{r}
x <- factor(c("yes", "yes", "no", "yes", "no"))
x
table(x)
unclass(x)
```

* Factors can set the order of levels using `levels` argument to `factor()`.

```{r}
x <- factor(c("yes", "yes", "no", "yes", "no"),
            levels = c("yes", "no"))
x
```

### Missing Values

* Missing Values are denoted by NA or NaN (mathematical operations) for underfined mathematical operations.
  * `is.na()` is used to test objects if they are `NA`
  * `is.nan()` is used to test for `NaN`
  * `NA` values have a class, so there are integer NA, character NA, etc.
  * A `NaN` value is also `NA` but the converse is not true

```{r}
x <- c(1, 2, NA, 10, 3)
is.na(x)
is.nan(x)
x <- c(1, 2, NaN, NA, 4)
is.na(x)
is.nan(x)
```

### Data Frames

* Data Frames are used to store tabular data
  * They are represented as a special type of list where every element of the list has to have the same length
  * Each element of the list can be thought of as a column and the length of each element of the list is the number of rows
  * Unlike matrices, data frames can store different class of objects in each column (just like lists), matrices must have every element be the same class
  * Data frames also have a special attribute called row.names
  * Data frames are usually created by calling `read.table()` or `read.csv()`
  * Can be converted to a matrix by calling `data.matrix()`

```{r}
x <- data.frame(foo = 1:4, bar = c(T, T, F, F))
x
nrow(x)
ncol(x)
```

### Names

* Names are for R objects, which is very useful for writing reable code and self-describing objects

```{r}
x <- 1:3
names(x)
names(x) <- c("foo", "bar", "norf")
x
names(x)
```

* Names for lists

```{r}
x <- list(a = 1, b = 2, c = 3)
x
```

* Names for matrices

```{r}
m <- matrix(1:4, nrow = 2, ncol = 2)
dimnames(m) <- list(c("a", "b"), c("c", "d"))
```
