Subsetting
==========

- `[`: returns an object of the same class, can be used to select more than one element
- `[[`: used to extract elements of a list or a data frame; it can only be used to extrat a single element and the class of the returned object will not necessarily be a list or data frame
- `$`: used to extract elemetns of a list or data frame by name; semantics are similar to that of `[[`

```{r}
x <- c("a", "b", "c", "d")
x[1]
x[1:4]
x[x > "a"]
u <- x > "a"
u
x[u] # select elments that are larger than "a"
```

#### subsetting a matrix

```{r}
x <- matrix(1:6, 2, 3)
x[1, 2] # returns a vector of length 1 rather than a 1x1 matrix
x[1, 2, drop=FALSE] # returns a 1x1 matrix

x[, 2] # omit indices
x[1, ] # returns a vector
x[1, , drop=FALSE] # returns a matrix
```

- `drop=FALSE`: returns a matrix rather than a vector

#### subsetting a list

```{r}
x <- list(foo = 1:4, bar = 0.6)
x[1] # returns a list containing the sequence
x[[1]] # returns a sequence
x$bar # returns a single value of "bar"
x[["bar"]] # same to x$bar
x["bar"] # return a list with "bar"
```

- The `[[` operator can be used with computed indices; `$` can only be used with liter names

```{r}
x <- list (foo = 1:4, bar = 0.6, baz = "hello")
name <- "foo"
x[[name]] # computed index for 'foo'
x$name # element 'name' does not exist
```

- The `[[` can take an integer sequence

```{r}
x <- list(a = list(10, 12, 14), b = c(3.14, 2.81))
x[[c(1, 3)]]
x[[1]][[3]]
x[[c(2,1)]] # same as x[[2]][[1]]
```

#### partial matching

```{r}
x <- list(aardvark = 1:5)
x$a
x[["a"]]
x[["a", exact = FALSE]] # same as x$a
```

#### removing `NA` values

```{r}
x <- c(1, 2, NA, 4, NA, 5)
bad <- is.na(x) # tell which elements are NA
x[!bad] #
```

```{r}
x <- c(1, 2, NA, 4, NA, 5)
y <- c("a", "b", NA, "d", NA, NA)
good <- complete.cases(x, y)
good
x[good]
y[good]
```
